Entropy Rain: Dilution and Compression of Thermals in Stratified Domains

### 2019-07-17 submission2/referee_response.txt

Dear Aimee,

We thank you and the anonymous referee for taking the time to consider and review our manuscript, "Entropy Rain: Dilution and Compression of Thermals in Stratified Domains".

In addition to the changes suggested in the referee report (and which we describe inline with the report below), we have made the following small changes to the manuscript as a result of discussions with colleagues:
1. We have computed a 3D, fully compressible simulation at n_rho = 3 and included its results in all relevant figures. We additionally recomputed our 3D, nrho = 2 simulation with a higher horizontal coefficient resolution.
2. We have added a line showing the sound speed to Figure 6, middle panel.
3. Throughout the manuscript, n_rho is now capital N_rho, to avoid confusion with n_ad.
4. Minor clarifications throughout the manuscript, all of which are bolded.

Best wishes,
Evan Anders, Daniel Lecoanet, and Ben Brown

----------------------------------------------------------
RESPONSE TO REFEREE REPORT
----------------------------------------------------------

""" 
The authors analyze the entropy rain scenario by developing an analytic
theory and performing simulations in the presence of negative buoyancy
of the downflow. This is a nice piece of work that should be published,
and I have only relatively minor suggestions for possible improvement.
"""

We thank the referee for taking the time to review our paper and for their useful suggestions to help improve the clarity of the manuscript.
We hope that the changes we have made to address the referees concerns, described below, are sufficient to clear up any points of confusion.

"""
1. The paper is arranged in such a way that subsequent simulation
results are used in quite some detail both in the introduction and in
the discussion of the model. If the authors want to keep it this way,
some additional clarifications should be given in several places (e.g.,
in the captions to Figs. 1-3, and in the text where they refer to these
figures.) Otherwise, details such as nrho=3, the size of the domain,
and other details of Fig.3 remain puzzling. Regarding Fig.2, one could
well imagine that it is just a sketch, but it is actually a simulation
result, right? If so, what are the times in the two panels? Another
option would be a rearrangement and to present the simulations first.
"""

We have elected to keep the overall presentation of the paper the same, but have added explicit callouts throughout the text to inform the reader that they are looking at simulation data in Figs 2 & 3.
See figure captions, the first sentences of section 2.1, and callouts to Fig 3 in section 2.3.

"""
2. Some quantities are not defined explicitly. This should be avoided,
even though the symbols follow standard convention (e.g., rho). It might
also be good to give a name to quantities such as t_off, when they are
first introduced. In the case of t_off, it appears first on page 3 in
Sect.2.3, and only at the end of Sect.2.4 it is called temporal offset.
"""

We have tried to explicitly define all parameters throughout the paper (rho in the sct. 1; u and g in 2.2; t_off, B_th, Gamma_th, m in 2.3, grad_ad in 2.4, and we redefine N_rho in 3.1 as its callout in 2.1 is far enough away from here that it can be forgotten).

"""
3. The expression detainment seems to be used in meteorology, but is
maybe new in astrophysics, where one talks in other contexts about
stripping or even ram pressure stripping. Maybe some comments could
be useful.
"""

We have added a few words to key the reader in on the concept of detrainment shortly before eqn 7.

"""
4. When B_th is first introduced in Sect.2.3, it is not clear to me in
which way it is different from just B that was used in the previous
section 2.2. Only in the second paragraph it seems to become clear
that this is because of detainment. Again, some text earlier on could
be helpful.
"""

We removed the first sentence of sct. 2.3 and have defined B_th and Gamma_th more clearly following eqns. 7 and 8, respectively.

"""
5. nabla_ad is sometimes put in parenthesis. In Eq.(10) one could imagine
that the authors would like to avoid an impression that it could be
an operator (in which case it could be put at the end of (z-L), but in
Eq.(15) it seems really unnecessary.
"""

We have removed these parenthesis and moved nabla_ad to the end of (z-L_z) in eqn. 10.

"""
6. In Eq.(14), should there not be a term in the viscosity expression
that is related to the derivative of the density?
"""

Due to our formulation of diffusivities (kappa = rho*chi = constant in space and time, mu = rho*nu = constant in space and time), the gradient terms are explicitly zero in our formulation.

"""
7. At the end of Sect.3.4, it becomes clear that periodic boundary
conditions are used in the vertical direction. It comes as a surprise
and it is unclear why this is done.
"""

We have added a paragraph explaining the reasoning behind this choice in section 3.4.

"""
8. In the caption to Fig.4, I'd add "(AN)" and "(FC)" after anelastic
and fully compressible. The differences are indeed quite small. Maybe
it would be good to emphasize that the author mean 2D Cartesian and
not 2D cylindrical, if that is indeed the case. 2D Cartesian implies a
very different width of vortices, so it is amazing that this does not
matter. Given that AN and FC are changed at the same time as 2D and 3D,
it is not clear what is caused by what. I agree, the differences are
small, but a comment might be useful. Is 3-D AN harder, or what?
"""

We have added (AN) and (FC) to the captions of figures 4 and 5.
Furthermore, we have added some additional text to 3.1, as it seems that it was unclear that the 3D simulations are in azimuthally symmetric cylindrical domains, while the 3D simulations are in cartesian domains.
Essentially, the 2D simulations are the limit in which a description of the thermal is easily derivable, and the 3D simulations are the "truth," and we wanted to compare those two.

"""
9. Except for general remarks in the introduction, nothing is said about
the enthalpy flux that is accomplished by one plume. It can obviously
not be modeled by a superadiabatic gradient, but maybe some quantitative
statements can be made regarding some of the assumptions made in B16.
"""

The focus of this work is entrainment, as suggested by the title, and entropy rain was merely the context in which we sought to apply this knowledge.
While interesting, descriptions of the flux carried by one of these thermals, especially propagated to the solar context, is outside of the remit of this manuscript.
We do not believe that this can be treated in a satisfying way in a short paragraph or two, and an aside of more length than that which does not address our original hypothesis seems out of place.

### 2019-07-17 submission2/referee_response.txt

Dear Dr. Norton,

We thank you and the anonymous referee for taking the time to consider and review our manuscript, "Entropy Rain: Dilution and Compression of Thermals in Stratified Domains".
We include below our changes to the manuscript, including inline responses to the full referee report.

@@

Best wishes,
Evan Anders, Daniel Lecoanet, and Ben Brown

----------------------------------------------------------
RESPONSE TO REFEREE REPORT
----------------------------------------------------------

In addition to the changes suggested in the referee report (which we describe inline below), we have made the following small changes to the manuscript as a result of discussions with colleagues:
1. We have computed a 3D, fully compressible simulation at n_rho = 3 and included its results in all relevant figures. We additionally recomputed our 3D, nrho = 2 simulation with a higher horizontal coefficient resolution.
2. We have added a line showing the sound speed to velocity panel of Figure 6.
3. Throughout the manuscript, n_rho is now capital N_rho, to avoid confusion with n_ad.
4. Minor clarifications throughout the manuscript, all of which are bolded.

@@

"""
5. nabla_ad is sometimes put in parenthesis. In Eq.(10) one could imagine
that the authors would like to avoid an impression that it could be
an operator (in which case it could be put at the end of (z-L), but in
Eq.(15) it seems really unnecessary.
"""

We have removed these parenthesis and moved nabla_ad to the end of (z-L_z) in eqn. 10.

"""
6. In Eq.(14), should there not be a term in the viscosity expression
that is related to the derivative of the density?
"""

Due to our formulation of diffusivities (kappa = rho*chi = constant in space and time, mu = rho*nu = constant in space and time), the gradients of density cancel with the gradients of nu and/or chi.

### 2019-07-17 tex/stratified_thermals.tex
incorporates Daniel's suggestions; does not yet put the flux discussion in (need to figure out how to handle PE + KE in addition to F_enth. Or do we ignore them?)

The 2D simulation domain is the direct product of a Fourier basis $(z \in [-L_z/4, L_z])$ and a Chebyshev basis $(r \in [0, L_r])$.
The boundary conditions are $\partial_r S_1 = w = (\grad\times\bm{u})_\phi = 0$ at $r = L_r$, and the regularity of the equations automatically impose $u = \partial_r(w) = \partial_r(S_1) = 0$ at $r = 0$.
The 3D simulation domain is the direct product of Fourier bases in the horizontal directions ($x, y \in [-L_r, L_r]$) and a Chebyshev basis ($z \in [0, L_z]$).
We impose impenetrable, stress free, fixed-temperature boundary conditions at the upper and lower boundaries.
In all of our simulations we specify $L_z = 20$ and $L_r = 5$.
We extend our 2D simulation domains to $z = -L_z/4$ because those simulations are vertically periodic.
This extension allows us to study the full transit of the thermal above $z = 0$ and terminate the simulation before it begins to wrap through the bottom of the periodic domain.

\textbf{
Dedalus currently allows for the direct specification of general boundaries in one direction.
In the case of our 2D simulations, this direction must be the radial direction in order to capture the singularity at $r = 0$.
While Dedalus can expand fields in terms of sine or cosine bases (as in \LJ), the mixed parities of terms such as the viscous heating term in the fully compressible or anelastic equations rules out a simple specification in these equation sets.
We have therefore chosen to use periodic Fourier bases in the vertical direction of our simulations, and end the simulations before the thermal can interact with those boundaries.
Remarkably, the results of these simulations vary only minimally from our 3D simulations, as we will show in the next section, and the vertical boundaries do not meaningfully affect the evolution of the thermal.
}

@@

\begin{deluxetable*}{c c c c c }
\tabletypesize{\footnotesize}
\caption{Table of simulation information
\label{table:simulation_info}
}
\tablehead{																																															
\colhead{$N_\rho$} & \colhead{nr or nx = ny} & \colhead{nz} & \colhead{$t_{\text{evolution}}$} & \colhead{safety}	}	
\startdata																																															
\multicolumn{5}{l}{\textbf{2D Anelastic Simulations}}\\
0.1 	& 		128			& 512			& 50 	&	0.6	\\
0.5 	& 		128			& 512			& 45 	&	0.6	\\
1	 	& 		128			& 512			& 41 	&	0.6	\\
2	 	& 		128			& 1024			& 34	&	0.4	\\
3	 	& 		192			& 1024			& 29	&	0.4	\\
4	 	& 		256			& 1024			& 26 	&	0.3	\\
5	 	& 		256			& 1536			& 25 	&	0.14	\\
6	 	& 		256			& 1536			& 23 	&	0.08	\\
\multicolumn{5}{l}{\textbf{3D Fully Compressible Simulations}}\\
0.1 	& 		256			& 512			& 50 		&	0.15	\\
0.5 	& 		256			& 512			& 45 		&	0.15	\\
1	 	& 		256			& 512			& 41	 	&	0.15	\\
2	 	& 		\textbf{384}			& 1024			& 34    	&	0.1	\\
\textbf{3}	 	& 		\textbf{512}			& \textbf{2048}			& \textbf{25}    	&	\textbf{0.15}	\\
\enddata																																															
\tablecomments{
Our 2D sims have spectral mode resolution of nr x nz, and our 3D sims have spectral mode resolution of nx x ny x nz.
We evolve our simulations from initial conditions until $t_{\text{evolution}}$ dimensionless freefall times have passed.
We use a CFL to determine the appropriate size of our timestep at each timestep in our simulation, and the CFL safety factor decreases with increasing stratification.
Our 2D sims use the RK443 timestepper and our 3D sims use the SBDF2 timestepper, resulting in different safety factor sizes in 2D and 3D.
}
\end{deluxetable*}

>>>>

The 2D simulation domain is \textbf{periodic in the z-direction with $z \in [-L_z/4, L_z]$ and the radial direction spans $r \in [0, L_r]$}.
The boundary conditions are $\partial_r S_1 = w = (\grad\times\bm{u})_\phi = 0$ at $r = L_r$, and the regularity of the equations automatically impose $u = \partial_r(w) = \partial_r(S_1) = 0$ at $r = 0$.
The 3D simulation domain is \textbf{periodic in the horizontal directions ($x, y \in [-L_r, L_r]$) and vertically spans $z \in [0, L_z]$}.
We impose impenetrable, stress free, fixed-temperature boundary conditions at the upper and lower boundaries.
In all of our simulations we specify $L_z = 20$ and $L_r = 5$.
We extend our 2D simulation domains to $z = -L_z/4$ because those simulations are vertically periodic.
This extension allows us to study the full transit of the thermal above $z = 0$ and terminate the simulation before it begins to wrap through the bottom of the periodic domain.

\textbf{
In our 2D simulations, we represent the radial direction with Chebyshev polynomials so that we can include geometric factors in ours equations and capture the singularity at $r = 0$.
We then assume the z direction is periodic to make the calculations more efficient, and we end the simulations before the thermal can interact with the vertical periodic boundaries.
Remarkably, the results of these simulations vary only minimally from our 3D simulations, as we will show in the next section, indicating that the vertical periodicity and the different side boundaries (cylinder vs. cube) do not influcence the thermal properties.
}

@@

\begin{deluxetable*}{c c c c c }
\tabletypesize{\footnotesize}
\caption{Table of simulation information
\label{table:simulation_info}
}
\tablehead{																																															
\colhead{$N_\rho$} & \colhead{nr or nx = ny} & \colhead{nz} & \colhead{$t_{\text{evolution}}$} & \colhead{safety}	}	
\startdata																																															
\multicolumn{5}{l}{\textbf{2D Anelastic Simulations}}\\
0.1 	& 		128			& 512			& 50 	&	0.6	\\
0.5 	& 		128			& 512			& 45 	&	0.6	\\
1	 	& 		128			& 512			& 41 	&	0.6	\\
2	 	& 		128			& 1024			& 34	&	0.4	\\
3	 	& 		192			& 1024			& 29	&	0.4	\\
4	 	& 		256			& 1024			& 26 	&	0.3	\\
5	 	& 		256			& 1536			& 25 	&	0.14	\\
6	 	& 		256			& 1536			& 23 	&	0.08	\\
\multicolumn{5}{l}{\textbf{3D Fully Compressible Simulations}}\\
0.1 	& 		256			& 512			& 50 		&	0.15	\\
0.5 	& 		256			& 512			& 45 		&	0.15	\\
1	 	& 		256			& 512			& 41	 	&	0.15	\\
2	 	& 		\textbf{384}			& 1024			& 34    	&	0.1	\\
\textbf{3}	 	& 		\textbf{384}			& \textbf{2048}			& \textbf{25}    	&	\textbf{0.15}	\\
\enddata																																															
\tablecomments{
Our 2D sims have spectral mode resolution of nr x nz, and our 3D sims have spectral mode resolution of nx x ny x nz.
We evolve our simulations from initial conditions until $t_{\text{evolution}}$ dimensionless freefall times have passed.
We use a CFL to determine the appropriate size of our timestep at each timestep in our simulation, and the CFL safety factor decreases with increasing stratification.
Our 2D sims use the RK443 timestepper and our 3D sims use the SBDF2 timestepper, resulting in different safety factor sizes in 2D and 3D.
}
\end{deluxetable*}
